### 2015-05-04 descriptive_analysis.Rmd
Added question code book and an Rmd for descriptive analysis.

---
title: "Untitled"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}

library(parsedate)
library(dplyr)
library(zoo)

(load("./JDAP 2_clean.rda"))
library(ggplot2)

d$PublicationDate <- parse_date(with(d, paste(month, year))) %>% as.yearqtr %>% as.character # as.yearqtr # as.yearmon
d$StartDate <- parse_date(d$StartDate) %>% as.yearqtr %>% as.character

x <- group_by(d, StartDate, Q=Q5) %>% dplyr::summarize(N.resp=length(StartDate)) %>% dplyr::mutate(N.total=sum(N.resp), P.resp=N.resp/N.total)
x$date <- parse_date(x$StartDate)
x <- x[order(x$date),]

View(x)
x$StartDate
subset(x, N.total > 1000)

ggplot(data=x) + aes(StartDate, P.resp, fill=Q)+geom_bar(stat="identity") #, order=Response, , fill=Q4

ggplot(d, aes(StartDate, Q4))+geom_bar(stat="identity")

qplot(Q4, data=d, geom="bar", fill=Q4)+facet_wrap(~StartDate)

library(HH)

head(d)

?likert

likert(t(apply(data, 2, table)))

summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.