### 2015-05-04 descriptive_analysis.Rmd
Added ordered logistic model results for the first two questions.

---
title: "Preliminary data analysis"
output: html_document
---

## Data pre-processing

Let's take a look at the numbers of responses by response date and publication date. Please keep in mind that the JDAP adoption dates used in the following analysis are approximate.

```{r, cache=TRUE, message=FALSE, echo=FALSE}

library(parsedate)
library(dplyr)
library(zoo)
library(ggplot2)
library(assertthat)

load("./data/JDAP 2_clean.rda")
load("./data/jdap_dates.rda")

# add the journal info to the main dataset
assert_that( all(d$journal %in% dates$journal) )
d <- left_join(d, dates, by="journal")

# format publication and response dates
d$PublicationDate <- parse_date(with(d, paste(month, year)))
d$ResponseDate <- parse_date(d$StartDate) %>% as.yearmon %>% as.POSIXct

d$ResponseDate_YM <- as.yearmon(d$ResponseDate) %>% as.character
d$ResponseDate_YQ <- as.yearqtr(d$ResponseDate) %>% as.character
d$ResponseDate_Y <- strftime(d$ResponseDate, "%Y") %>% as.character

d$PublicationDate_YM <- as.yearmon(d$PublicationDate) %>% as.character
d$PublicationDate_YQ <- as.yearqtr(d$PublicationDate) %>% as.character
d$PublicationDate_Y  <- strftime(d$PublicationDate, "%Y") %>% as.character

# determine time since JDAP introduction in the journal where the article was published
d$DaysPolicyToPublication <- difftime(d$PublicationDate, parse_date(d$policy_start), units="days")
d$DaysPolicyToResponse <- difftime(d$ResponseDate, parse_date(d$policy_start), units="days")

d$PolicyAdoptedAtPublication <- ifelse(d$DaysPolicyToPublication>0, "After JDAP", "Before JDAP")
d$PolicyAdoptedAtResponse <- ifelse(d$DaysPolicyToResponse>0, "After JDAP", "Before JDAP")
d$PolicyAdoptedAtPublication[is.na(d$PolicyAdoptedAtPublication)] <- "No JDAP"
d$PolicyAdoptedAtResponse[is.na(d$PolicyAdoptedAtResponse)] <- "No JDAP"
d$PolicyBinary <- ifelse(d$policy=="Required", "JDAP", "No JDAP")

# number of responses by response date and journal policy in 2015
group_by(d, Date=ResponseDate_YQ, policy) %>% dplyr::summarize(N=length(StartDate)) %>% ggplot(aes(Date, N, color=policy, group=policy))+geom_point()+geom_line()+theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Number of participants by response date.")

# number of responses by publication date and journal policy in 2015
group_by(d, Date=PublicationDate_YQ, policy) %>% dplyr::summarize(N=length(PublicationDate)) %>% ggplot(aes(Date, N, color=policy, group=policy))+geom_point()+geom_line()+theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Number of participants by publication date.")

# It seems to make more sense to focus on the publication date for plotting, until it is clear why there are hardly any responses from the second half of 2013.
```

## Critical questions

In the following, I will present an overview of the responses to questions 1 and 2. The questions asked about the strength of agreement/disagreement with the following statements:

* _Question 1_ It is the community norm in my field to share, with qualified researchers, datasets that support the results of peer-reviewed research articles. Include data sharing that occurs by any mechanism (e.g. emailing datasets upon request, posting datasets to lab websites, including datasets in journal supplementary information or data repositories).
* _Question 2_ It is the community norm in my field to publicly archive all supporting datasets online, for use by anyone for any purpose, upon publication of a peer-reviewed research article.

## Attitudes towards data sharing and archiving by journal type

The following plots show the responses regarding the attitudes towards data sharing and archiving by whether or not the journal has adopted a JDAP-equivalent data archiving policy by now (May 2015).

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
source("src/multiplot.R")

x <- group_by(d, PolicyBinary, Response=Q4) %>% dplyr::summarize(N.resp=length(StartDate)) %>% filter(!is.na(Response)) %>% dplyr::mutate(N.total=sum(N.resp), P.resp=N.resp/N.total, Policy=sprintf("%s\n(N=%d)", PolicyBinary, N.total))
x <- x[order(x$Response, decreasing=T),]
p1 <- ggplot(data=x) + aes(Policy, P.resp*100, fill=Response, alpha=Response)+geom_bar(stat="identity")+ggtitle("'Data sharing is a community norm.'")+scale_fill_manual(values=c(rep("blue",3),"grey",rep("red",3)))+scale_alpha_manual(values=c(1,.75,.5,1,.5,.75,1))+theme_bw()+xlab("Current policy")+ylab("Percentage")

x <- group_by(d, PolicyBinary, Response=Q5) %>% dplyr::summarize(N.resp=length(StartDate)) %>% filter(!is.na(Response)) %>% dplyr::mutate(N.total=sum(N.resp), P.resp=N.resp/N.total, Policy=sprintf("%s\n(N=%d)", PolicyBinary, N.total))
x <- x[order(x$Response, decreasing=T),]
p2 <- ggplot(data=x) + aes(PolicyBinary, P.resp*100, fill=Response, alpha=Response)+geom_bar(stat="identity")+ggtitle("'Data archiving is a community norm.'")+scale_fill_manual(values=c(rep("blue",3),"grey",rep("red",3)))+scale_alpha_manual(values=c(1,.75,.5,1,.5,.75,1))+theme_bw()+xlab("Current policy")+ylab("Percentage")

p1

p2

```

The same plots side by side. Clearly, data sharing is considered more of a community standard than data archiving.
```{r, echo=FALSE, fig.width=10}

multiplot(p1, p2, cols=2)
```

## Attitudes towards data sharing and archiving by journal type and journal JDAP adoption date

The following plots show the responses regarding the attitudes towards data sharing and archiving for participants who (i) published in journals which have note adopted the JDAP so far ('No JDAP'), (ii) published in journals which had not _yet_ adopted the JDAP before publication ('Before JDAP'), and (iii) journals which had adopted the JDAP before publication ('After JDAP).

```{r, echo=FALSE}
x <- group_by(d, PolicyAdoptedAtPublication, Response=Q4) %>% dplyr::summarize(N.resp=length(StartDate)) %>% filter(!is.na(Response)) %>% dplyr::mutate(N.total=sum(N.resp), P.resp=N.resp/N.total, Policy=sprintf("%s\n(N=%d)", PolicyAdoptedAtPublication, N.total))
x <- x[order(x$Response, decreasing=T),]
p1 <- ggplot(data=x) + aes(Policy, P.resp*100, fill=Response, alpha=Response)+geom_bar(stat="identity")+ggtitle("'Data sharing is a community norm.'")+scale_fill_manual(values=c(rep("blue",3),"grey",rep("red",3)))+scale_alpha_manual(values=c(1,.75,.5,1,.5,.75,1))+theme_bw()+xlab("Policy at the time of publication")+ylab("Percentage")

x <- group_by(d, PolicyAdoptedAtPublication, Response=Q5) %>% dplyr::summarize(N.resp=length(StartDate)) %>% filter(!is.na(Response)) %>% dplyr::mutate(N.total=sum(N.resp), P.resp=N.resp/N.total, Policy=sprintf("%s\n(N=%d)", PolicyAdoptedAtPublication, N.total))
x <- x[order(x$Response, decreasing=T),]
p2 <- ggplot(data=x) + aes(Policy, P.resp*100, fill=Response, alpha=Response)+geom_bar(stat="identity")+ggtitle("'Data archiving is a community norm.'")+scale_fill_manual(values=c(rep("blue",3),"grey",rep("red",3)))+scale_alpha_manual(values=c(1,.75,.5,1,.5,.75,1))+theme_bw()+xlab("Policy at the time of publication")+ylab("Percentage")

p1

p2

```

The same plots side by side. Clearly, data sharing is considered more of a community standard than data archiving.
```{r, echo=FALSE, fig.width=10}

multiplot(p1, p2, cols=2)
```

### Inferential statistics

```{r, message=FALSE, eval=TRUE, echo=FALSE}

library(MASS)
library(zoo)
library(dplyr)
library(knitr)

d$cAfterPolicyChange <- ifelse(d$PolicyAdoptedAtPublication=="After JDAP", 1, 0)
d$cNonJDAPJournal <- ifelse(d$PolicyAdoptedAtPublication=="No JDAP", 1, 0)
d$cDate <- as.double(as.yearmon(d$PublicationDate)) %>% scale(., scale=F)

d$respQ4 <- ordered(d$Q4, levels=rev(levels(d$Q4)))
d$respQ5 <- ordered(d$Q5, levels=rev(levels(d$Q5)))

m4 <- polr(respQ4 ~ 1 + cAfterPolicyChange + cNonJDAPJournal + cDate, data=d , Hess=TRUE)
# summary(m4)
m4.justchange <- polr(respQ4 ~ 1 + cAfterPolicyChange, data=d , Hess=TRUE)
m4.nochange <- polr(respQ4 ~ 1 + cNonJDAPJournal + cDate, data=d , Hess=TRUE)
suppressMessages( ci4 <- confint(m4) )
table4 <- exp(cbind(OR = coef(m4), ci4)) %>% cbind(., t=coef(summary(m4))[1:nrow(.),'t value'])
#table4 <- cbind(lOR = coef(m4), ci4) %>% cbind(., t=coef(summary(m4))[1:nrow(.),'t value'])

#anova(m4, m4.justchange, m4.nochange)

m5 <- polr(respQ5 ~ 1  + cAfterPolicyChange + cNonJDAPJournal + cDate, data=d, Hess=TRUE)
m5.justchange <- polr(Q5 ~ 1 + cAfterPolicyChange , data=d, Hess=TRUE)
m5.nochange <- polr(Q5 ~ 1 + cNonJDAPJournal + cDate, data=d, Hess=TRUE)
# summary(m5)
suppressMessages( ci5 <- confint(m5) )
table5 <- exp(cbind(OR = coef(m5), ci5)) %>% cbind(., t=coef(summary(m5))[1:nrow(.),'t value'])

#anova(m5, m5.justchange, m5.nochange)
```

The potential effect of JDAP adoption on authors' attitudes is might be confounded by two other factors. Firstly, groups of researchers publishing in journals which are not planning to adopt the JDAP (or at least have not done so yet) may inherently differ from groups publishing in journals which were open to adopting the JDAP. Secondly, attitudes regarding data archiving may improve over time.

In order to test for an effect of policy change on attitudes while controlling for potential confounding effects, we used an ordinal logistic regression model with three predictors: (i) journal type _(cNonJDAPJournal)_, (ii) (centered) publication date in years _(cDate)_, and (iii) a variable indicating whether the article was published after the the journal adopted a data archiving policy _(cAfterPolicyChange)_.

The following table shows the estimated coefficients and 95\%-confidence intervals (transformed to odds ratios) for responses to *question 1*. Effects are statistically significant at the $\alpha=0.05$ level when the transformed confidence intervals do not contain $1$. Confidence intervals were obtained by profile likelihood. According to the model estimates, the introduction of the JDAP data archiving policy increases the odds of a positive response regarding *data sharing* by factor $1.36$ (CI=$[1.16; 1.60]$), while the remaining two predictors do not show any statistically significant effects.
```{r, message=FALSE, eval=TRUE, echo=FALSE}
options(digits=2)
table4
```

The following table shows the estimated coefficients and 95\%-confidence intervals (transformed to odds ratios) for responses to *question 2*. According to the model estimates, the introduction of the JDAP data archiving policy increases the odds of a positive response regarding *data archiving* by factor $1.57$ (CI=$[1.33; 1.90]$), while the remaining two predictors do not show any statistically significant effects.

```{r, message=FALSE, eval=TRUE, echo=FALSE}
options(digits=2)
table5
```